\huge Modeling Intervention Strategies for United States TB Control

### 2013-07-29 in_progress/presentations/poster/reuMHC2013NoSIRIntro.tex
with new text and visual stuff

\begin{frame}
  \begin{columns}
    \begin{column}{.3\textwidth}
      \vspace{-2em}
      \begin{block}{Introduction}
        \begin{itemize}
          \item Epidemiological models describe disease outbreaks.
          \item Compartmental Differential Equation models are common and
            rigorous epidemiological models.
          \item The Hill Model is a complex compartmental model of tuberculosis
            (TB) in the United States (US).
        \end{itemize}
       % Epidemiological models offer insight into the structure of disease
       % outbreaks and the merits of various interventions. The most common
       % epidemiological models are compartmental differential equation models,
       % in which the population is dispersed between a variety of distinct
       % compartments, through which they move according to a system of
       % Differential Equations. In order to model tuberculosis (TB) in the
       % United States (US), Hill, Becerra, and Castro designed a complex
       % compartmental model called the hill model.
      \end{block}
      \begin{block}{The Basic Hill Model}
        \vspace{-2em}
        \begin{block}{}
          \begin{column}{.5\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=12cm,width=.9\textwidth]{HillModelFlowChart}
              \end{center}
              \caption{A flow chart representing the compartments of the Hill
                       Model.}
              \label{fig:hillFlow}
            \end{figure}
          \end{column}
          \begin{column}{.43\textwidth}
            Populations:
            \begin{itemize}
              \item US Born Individuals (USB) 
              \item Foreign Born Individuals (FB)
            \end{itemize}
            Individuals also leave the model due to natural death.
          \end{column}
          \begin{figure}[h]
            \begin{center}
              \includegraphics[height=25cm,width=\textwidth]{incPlotSourced2}
            \end{center}
            \caption{The source population of US TB incindecence}
            \label{fig:incPlotSourced}
          \end{figure}
          \vspace{-1em}
          \begin{figure}[h]
            \begin{center}
              \includegraphics[height=15cm,width=\textwidth]{incPlotTotal}
            \end{center}
            \caption{New cases per year of various types of TB in the US.}
            \label{fig:incPlotTotal}
          \end{figure}
        \end{block}
      \end{block}
    \end{column}

    \begin{column}{.3\textwidth}
      \vspace{-1.5em}
      \begin{block}{Analyzing US TB Reduction Strategies}
        \begin{itemize}
          \item Implemented in \texttt{R}, with various numerical DE solvers.
          \item Tracks US Health Care System (HCS) cost.
          \item Tracks statistics about various health states.
        \end{itemize}
        \begin{block}{Intervention Analysis}
          \vspace{-1.6em}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=10cm,width=\textwidth]{incidencePlotRedEnLTBI}
              \end{center}
              \caption{Incidence/million in USB, FB, and total populations,
                       given 0\%, 50\%, 75\%, or 100\% treatment of incoming
                       LTBI.}
              \label{fig:redEnLTBI_incidence}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=10cm,width=\textwidth]{incLTBItrmtIncGrouped}
              \end{center}
              \caption{Incidence/million in USB, FB, and total populations,
                       given 0\%, 100\%, or 300\% LTBI treatment increase.}
              \label{fig:incLTBItrmt_incidence}
            \end{figure}
          \end{column}
        \end{block}
      \end{block}
      \begin{block}{Economic Modeling}
        \begin{itemize}
          \item Tracks treatment costs for various disease states
          \item Estimates implementation cost of intervention
        \end{itemize}
        \begin{block}{}
          \vspace{-4em}
          \begin{column}{.66\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth,height=13cm]{EnLTBIRedGroupCost.pdf}
              \end{center}
              \caption{Cumulative implementation costs, US HCS savings, and net
                       US costs of LTBI arrival cure rates. Cost/case cured was
                       \$600, \$800, and \$1000 for 50\%, 75\%, and 100\%
                       cured.}
              \label{fig:redEnLTBI_costs}
            \end{figure}
          \end{column}
          \begin{column}{.26\textwidth}
            Base HCS Costs:
            \begin{description}
              \item[Active TB:]\hfill \\ 
                \$14,014.90
              \item[LTBI:]\hfill \\ 
                \$403.45
            \end{description}
          \end{column}
        \end{block}
      \end{block}
    \end{column}
    \begin{column}{.3\textwidth}
      \vspace{-.5em}
      \begin{block}{An Agent Based Implementation}
        Agent based models capture disease dynamics on the individual level, and
        reflect stochasticity and granularity lost in compartmental models.
        Agent based counterparts to the Hill model were implemented in Netlogo
        and \texttt{c++}.
        \begin{block}{}
          \vspace{-2em}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{NLHMinc}
              \end{center}
              \caption{Incidence/million for R and NetLogo models (12 runs, $\Delta t$ = 0.1, popConst = 100)}
              \label{fig:NLHMinc}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{finalRunSmall}
              \end{center}
              \caption{Incidence/million for R and C++ models (2100 runs, $\Delta t$ = 0.01, popConst = 1)}
              \label{fig:finalRun}
            \end{figure}
          \end{column}
        \end{block}
      \end{block}

      \begin{block}{Stochastic Models as a Measure of Variability}
        The stochastic model provides data on the variability of the results
        of the deterministic model.
        \begin{block}{}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{IN0dist}
              \end{center}
              \caption{Distribution of USB Incidence (C++) with fitted Normal curve}
              \label{fig:IN0dist}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{IN1dist}
              \end{center}
              \caption{Distribution of FB Incidence (C++) with fitted Normal curve}
              \label{fig:IN1dist}
            \end{figure}
          \end{column}
        \end{block}
      \end{block}

      \begin{block}{Future Extensions}
        \begin{itemize}
          \item Including contact structure
          \item Multi-drug resistant TB
          \item HIV
        \end{itemize}
      \end{block}

>>>>

\begin{frame}
  \begin{columns}
    \begin{column}{.3\textwidth}
      \vspace{-2em}
      \begin{block}{Introduction}
        Epidemiological models offer insight into the structure of disease outbreaks and the merits of various interventions.  Compartmental differential equation models are a common model in which populations move between various health states, or compartments, according to predetermined rates.  This work is an extension of the Hill Model, a complex compartmental model of tuberculosis (TB) in the United States.
       % Epidemiological models offer insight into the structure of disease
       % outbreaks and the merits of various interventions. The most common
       % epidemiological models are compartmental differential equation models,
       % in which the population is dispersed between a variety of distinct
       % compartments, through which they move according to a system of
       % Differential Equations. In order to model tuberculosis (TB) in the
       % United States (US), Hill, Becerra, and Castro designed a complex
       % compartmental model called the hill model.
      \end{block}
      \begin{block}{The Basic Hill Model}
        \vspace{-2em}
        \begin{block}{}
          \begin{column}{.5\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=12cm,width=.9\textwidth]{HillModelFlowChart}
              \end{center}
              \caption{A flow chart representing the compartments of the Hill Model.}
              \label{fig:hillFlow}
            \end{figure}
          \end{column}
          \begin{column}{.43\textwidth}
            Populations:
            \begin{itemize}
              \item US Born (USB) 
              \item Foreign Born (FB)
            \end{itemize}
            Individuals also leave the model due to natural death.
          \end{column}
          \begin{figure}[h]
            \begin{center}
              \includegraphics[height=20cm] {newfig}%{incPlotSourced2}
            \end{center}
            \caption{The source population of US TB incidence}
            \label{fig:incPlotSourced}
          \end{figure}
          \vspace{-1em}
%          \begin{figure}[h]
%            \begin{center}
%              \includegraphics[height=15cm]{incPlotTotal}
%            \end{center}
%            \caption{New cases per year of various types of TB in the US.}
%            \label{fig:incPlotTotal}
%          \end{figure}
        \end{block}
      \end{block}
    \end{column}

    \begin{column}{.3\textwidth}
      \vspace{-1.5em}
      \begin{block}{Analyzing US TB Reduction Strategies}
        \begin{itemize}
          \item Implemented in \texttt{R}, with various numerical DE solvers
          \item Tracks US Health Care System (HCS) cost
          \item Tracks statistics about various health states
        \end{itemize}
        \begin{block}{Intervention Analysis}
          \vspace{-1.6em}
          %\begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=10cm,width=\textwidth]{incidencePlotRedEnLTBI}
              \end{center}
              \caption{Incidence/million in USB, FB, and total populations,
                       given 0\%, 50\%, 75\%, or 100\% treatment of incoming
                       LTBI.}
              \label{fig:redEnLTBI_incidence}
            \end{figure}
          %\end{column}
          %\begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[height=10cm,width=\textwidth]{incLTBItrmtIncGrouped}
              \end{center}
              \caption{Incidence/million in USB, FB, and total populations,
                       given 0\%, 100\%, or 300\% LTBI treatment increase.}
              \label{fig:incLTBItrmt_incidence}
            \end{figure}
          %\end{column}
        \end{block}
      \end{block}
      \begin{block}{Economic Modeling}
        \begin{itemize}
          \item Tracks treatment costs for various disease states
          \item Estimates implementation cost of intervention
        \end{itemize}
        \begin{block}{}
          \vspace{-4em}
          \begin{column}{.66\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth,height=13cm]{EnLTBIRedGroupCost.pdf}
              \end{center}
              \caption{Cumulative implementation costs, US HCS savings, and net
                       US costs of LTBI arrival cure rates. Cost/case cured was
                       \$600, \$800, and \$1000 for 50\%, 75\%, and 100\%
                       cured.}
              \label{fig:redEnLTBI_costs}
            \end{figure}
          \end{column}
          \begin{column}{.26\textwidth}
            Base HCS Costs:
            \begin{description}
              \item[Active TB:]\hfill \\ 
                \$14,014.90
              \item[LTBI:]\hfill \\ 
                \$403.45
            \end{description}
          \end{column}
        \end{block}
      \end{block}
    \end{column}
    \begin{column}{.3\textwidth}
      \vspace{-.5em}
      \begin{block}{An Agent-Based Implementation}
        Agent-based models capture disease dynamics on the individual level, and
        reflect stochasticity and granularity lost in compartmental models.
        Agent-based counterparts to the Hill model were implemented in Netlogo
        and \texttt{C++}.
        \begin{block}{}
          \vspace{-2em}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{NLHMinc}
              \end{center}
              \caption{Incidence/million for R and NetLogo models (12 runs, $\Delta t$ = 0.1, popConst = 100)}
              \label{fig:NLHMinc}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{finalRunSmall}
              \end{center}
              \caption{Incidence/million for R and C++ models (2100 runs, $\Delta t$ = 0.01, popConst = 1)}
              \label{fig:finalRun}
            \end{figure}
          \end{column}
        \end{block}
      \end{block}

      \begin{block}{Stochastic Models as a Measure of Variability}
        \begin{block}{}
          \vspace{-2em}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{IN0dist}
              \end{center}
              \caption{Distribution of USB Incidence (C++) with fitted Normal curve}
              \label{fig:IN0dist}
            \end{figure}
          \end{column}
          \begin{column}{.45\textwidth}
            \begin{figure}[h]
              \begin{center}
                \includegraphics[width=\textwidth]{IN1dist}
              \end{center}
              \caption{Distribution of FB Incidence (C++) with fitted Normal curve}
              \label{fig:IN1dist}
            \end{figure}
          \end{column}
        \end{block}
      \end{block}

      \begin{block}{References}
        Hill, A. N., Becerra, J. E., \& Castro, K. G. (2012). Modelling tuberculosis trends in the USA. Epidemiology and infection, 140(10), 1862.
      \end{block}
